---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# {matchwith}

## What is this package?
match_with() simulates pattern matching that is used in many functional
programming language such as 'case of' in Haskell and 'match with' in OCaml.

## Installation
``` r
# install.packages("devtools")
devtools::install_github("tobcap/matchwith")
library("matchwith")
```
## Supports
* Constant pattern
* Cons pattern
* Tuple pattern with matching symbols
* Wildcard pattern
* Guard clauses (not completed implemented)

```{r load_packages, include=FALSE}
library(matchwith)
```

## Usage
```
# Syntax
f <- function(expr) {
  match_with(expr
  , pattern_1 -> res_1
  , pattern_2 -> res_2
             ...
  , pattern_n -> res_n
  )
}
```

# Const pattern
`pattern_i` represents numeric literal, charater literal, or NULL.
When x is matched with `pattern_i`, `res_i`
(the right side of expression) is evalueted in the parent.frame .
In the exampe below, when x is 0 or 1, the right side expression is evalated
and returned. when x is more than 2, wildcard simbol is matched and
`fib(x-1) + fib(x-2)` is evaluated and returned.
```{r}
fib <- function(x)
  match_with(x
  , 0 -> 0
  , 1 -> 1
  , . -> fib(x - 1) + fib(x - 2)
  )
fib(10)
```

# Cons pattern
`::` is used for DSL porpose that is representing constructor of head and tail.
when pattern is `y::ys`, the first object of `x` is assigned to `y` and rest
of that is assigned to `ys`.
```{r}
sum1 <- function(x)
  match_with(x
  , integer(0) -> 0L
  , y::ys -> y + sum1(ys)
  )
sum1(1:5)
```
```{r}
sum2 <- function(x, acc = 0L)
  match_with(x
  , integer(0) -> acc
  , y::ys -> sum2(ys, acc + y)
  )
sum2(1:5)
```

# Tuple pattern with matching
When `pattern_i` starts with `list` and includes new symbols as
list's arguments, the coresponding values of `x` which must have
the same language structure will be assigned to the new symbols, 
and `res_i` is evaluated with using such symbol-value-associated-list.

```{r}
ex1 <- function(x)
  match_with(x
  , list(a, b) -> 10 * a + b
  , list(a, b, c) -> 100 * a + 10 * b + c
  , . -> x
  )
ex1(list(1,2))
ex1(list(1,2,3))
str(ex1(list(1,2,3,4)))
```

Wildcard inside `list` of `pattern_i` can be used.
```{r}
fizzbuzz1 <- function(x)
  match_with(list(x %% 3, x %% 5)
  , list(0, 0) -> "fizzbuzz"
  , list(0, .) -> "fizz"
  , list(., 0) -> "buzz"
  , . -> as.character(x)
  )
sapply(1:30, fizzbuzz1)
```

Other than `list`, `x` is compared with `pattern_i` and if it matches, `res_i`
is evaluated.
```{r}
ex2 <- function(x)
  match_with(x
  , 1:3 -> 2 * x
  , 1:5 -> 3 * x
  , . -> 100 * x
  )
ex2(1:3)
ex2(1:5)
ex2(1:4)
ex2(c(1,2,3)) # 1:3 is integer class, c(1,2,3) is numeric class
ex2(c(1L,2L,3L))
```

# Guard clauses
when `pattern_i` uses `Compare` family (or `!`, `any`, `all`,
`identity`, `isTRUE`) or a function whose name starts with `is.` as top of
node of language object, it is regarded as guard clauses, and when the
result of evaluating `pattern_i` is TRUE, `res_i` will be evaluated
and returned.
```{r}
ex3 <- function(x)
  match_with(x
  , x > 5 -> x * 100
  , x < 3 -> x * 10
  , . -> -x
  )
sapply(1:10, ex3)

# isTRUE(x > 5) -> FALSE, isTRUE(x < 2) -> FALSE, 
# so only wildcard is matched and returns `-x`
ex3(1:10) 
```

  
